---
title: "Choosing Starting Estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing Starting Estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: starting-estimates.bib
link-citations: yes
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(collapse = TRUE, comment = "#>", cache = TRUE)
```

DW-NOMINATE uses a local optimization algorithm to get an approximate maximum
likelihood estimate of legislator ideal points. However, the likelihood function
is complex and can contain local maxima that aren't the global
maximum. Therefore it's important to provide starting estimates close to the
global maximum in order for DW-NOMINATE to find the correct results.

There's currently no single best approach for creating starting estimates. The
choice of approach depends on properties of the data.

# wnominate and oc

The `wnominate` and `oc` packages in R can create static ideal point estimates
using W-NOMINATE and optimal classification methods, respectively. Neither
package is designed to work with multiple legislative sessions, but they can be
used for this purpose by combining the data from multiple sessions into a single
`rollcall` object.

The downside is that both W-NOMINATE and optimal classification can take much
longer to run than DW-NOMINATE when there are many sessions. Of the two, `oc` is
typically faster.

```{r static, eval=FALSE}
library(dwnominate)
library(oc)
data(nhsenate)

# combine NH Senate rollcall objects
combined_nhsenate <- dwnominate:::merge.rollcall(rc_list = nhsenate)
# get optimal classification scores for the combined rollcalls
starts <- oc(combined_nhsenate, polarity = c(1, 1))
```

# common_space

Another strategy for deriving static scores is to calculate scores for each
session, and combine the results into a single overall score. That's the
approach used by the common space method.

```{r common-space, eval=FALSE}
library(dwnominate)
library(wnominate)
data(nhsenate)

# run W-NOMINATE on every session of the NH Senate
wnom_list <- lapply(nhsenate, wnominate, polarity = c(1, 1))

starts <- common_space(wnom_list)
```

# Manual strategies

@carroll_measuring_2009 [p. 265] describes the original process for creating the
starting values:

> DW-NOMINATE like its predecessor D-NOMINATE does not generate its own
> parameter starting values because of the sheer size of the roll call data set
> being analyzed. Good starting values are crucial to reliably performing this
> complex nonlinear estimation. In their development of D-NOMINATE, Poole and
> Rosenthal experimented for over 2 years until they had a satisfactory solution
> to the problem. Essentially, they began with what they thought was a sensible
> set of starting values and studied the estimated legislator coordinates by
> turning them into computer animations that could be played on a VHS
> videorecorder. In these animations, letter tokens and colors for the estimated
> legislator coordinates were used to identify parties and regions of the United
> States. Poole and Rosenthal then relied upon their knowledge of American
> political history to see if the output made visual sense. If they saw
> anomalies, they adjusted the output coordinates to correct for the anomalies
> and then used these adjusted coordinates as new starting values. At each step
> of this process, the fit of the model increased but it was a relatively
> arduous task because the supercomputers of the 1980s were limited in their
> capabilities and the animations took about a week to make on the equipment at
> the Pittsburgh Supercomputer Center.
>
> The upshot of what one of us has described as "Poole and Rosenthal as the
> outer loop of the estimation" is that the final coordinates from D-NOMINATE
> that were released to the larger research community in 1989–90 were about as
> close to the global maximum of the (constrained) likelihood function shown in
> equation (10) as was practicable given the computer resources of the time.
>
> Consequently, when DW-NOMINATE was developed in 1996 it simply used for starts
> the D-NOMINATE coordinates along with patched-in W-NOMINATE coordinates for
> Congresses 100–105.

@poole_spatial_2005 [p. 140] describes a simpler approach, piecing together
common space scores estimated on subsets of the data:

> I developed the CSS method in 1982, and Howard Rosenthal and I used it along
> with a simple metric scaling method (Poole 1984, 1990; Poole and Rosenthal
> 1997) to get starting coordinates in one and two dimensions for
> DW-NOMINATE. We used common-space coordinates for stable periods and then
> pieced together all the sets of starting coordinates for Congresses 1
> to 100.

# References
